---
title: "ESM244 Lab 4"
author: "Sara Orofino"
date: "1/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

0. Attach packages 

```{r, message=FALSE}

library(tidyverse)
library(plm)
library(Ecdat)
library(lmtest)
library(car)

```

1. Part 1: Nonlinear least squares (logistic growth of bacteria)

```{r}


cell_growth <- read_csv("CellGrowth.csv")

# Next look at data in ggplot: 

ggplot(cell_growth, aes(x=Time, y=CellCount)) +
  geom_point()

# To use nonlinear logistic growth you have to specify the model and the beginning estimates 
# Logistic growth equation Nt = A/(1+Be^(-rt))

# Find estimates for logistic growth (A, B and r):
# Estimate for carrying capacity (A) - 3700
# If N0 is 2000, what is b? 
   ## B = (K -N0)/N0 ~ 0.85
# r = 0.1121

BEstimate <- (3700-2000)/2000


# Range where exponetial growth (r) is dominate is about 0-3 hours
# Plot the natural log of cell count vs. time will give you r (linear relationship between ln(y) and t with r as the slope of that line)

# Now, we're going to estimate r (growth rate):

lag_phase <- cell_growth %>% 
  filter(Time < 4)
  
ggplot(lag_phase, aes(x= Time, y = log(CellCount))) +
  geom_point()
  
lm(log(CellCount) ~ Time, data = lag_phase)

# r = 0.1121 -- this is using the coefficient for Time in the lm model 

```

Find A, B, and r using Gauss-Newton method for iterative convergence (nonlinear least squares):
Use the nls() function:


```{r}

# First input your model function (in this case logistic growth), use start = list() to give the initial estimates for the parameters, give dataframe

cell_fit <- nls(CellCount ~ A/(1+B*exp(-r*Time)),
                start = list(A = 3700, B = 0.85, r = 0.1121),
                data = cell_growth, 
                trace = TRUE)

# Use trace to see all the iterations r ran in order to find convergence (minimized least squares for the parameter values)

cell_fit


# Store actual parameters for use in the model - use the coefficients from the model in case the model changes
A <- coef(cell_fit)[1]
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]

# Create a new sequence of time values (time_seq) to use to make predictions 

time_seq <- seq(0,20, length=100) # use length for the total number of observations in that sequence

# Plug that new sequence into the model with the parameters for A, B and r that we found:

cell_pred <- A/(1+B*exp(-r*time_seq))

# Bind that with time sequence data using data.frame():
pred_df <- data.frame(time_seq, cell_pred)

# Create a graph with the original data and our model predictions: 

ggplot(cell_growth, aes(x = Time, y = CellCount)) +
  geom_point(color = "blue", size = 3) +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred), color = "orange", size = 1) + 
  theme_light()

# Note keep the color and size outside of aes() because they don't reference a specific variable and if you put them inside the aes() ggplot will try to make a legend for it 

```


Panel Regression - First Shot 

State: our entity variable 
Year: our time variable 
packpc: our dependent variable 
avgprs: our predictor variable of interest 

```{r}

cigs_panel <- Cigarette %>% 
  select(state, year, packpc, avgprs)


# Look at how pricing is affecting purchasing:

ggplot(cigs_panel, aes(x = avgprs, y = packpc, group = state)) +
  geom_point(aes(color = state)) + 
  geom_smooth(method = "lm", aes(color=state), se = FALSE, size = 0.2)

# Overall general trends are decreasing for every state, average price looks to be affecting purchases in a pretty similar way across each state 

# Just to see look at how years are affecting purchasing 

ggplot(cigs_panel, aes(x = year, y = packpc), group = state) +
  geom_point(aes(color=state)) +
  geom_smooth(method = "lm", aes(color=state), se = FALSE, size = 0.2)

# Also just general trends over time to reduce purchases of cigarettes (like campaigns or restrictions), probably should consider time fixed effects
```

Entity Fixed Effects model:

```{r}
# model = "within" for entity fixed effects model

cigs_model <- plm(packpc ~ avgprs,
                  data = cigs_panel,
                  index = c("state", "year"),
                  model = "within")

cigs_model

# Coefficient -0.32545
# Interpretation: When we account for entity fixed affects, for every $1 in price we would expect to see a decrease in per capita cigarette purchases of 0.32545.


# Robust Standard Error calculation: 

coeftest(cigs_model, vcov. = vcovHC(cigs_model, type = "HC1"))

# Std. Error metric gives you the robust standard error, notice that the estimate is slightly more conservative than the model prediction coefficient.
```

Add time fixed effects (such as):
* Marketing campaigns 
* Regulations for packaging and distribution 

```{r}
# Only need to add effect = "twoways" to consider time fixed effects in the model in addition to entity fixed effects 

et_cigs_model <- plm(packpc ~ avgprs,
                     data = cigs_panel,
                     index = c("state", "year"),
                     model = "within", 
                     effect = "twoways")
et_cigs_model

# To compare the two models and see if it got better with the time fixed effect
# H0: Model is not improved by adding time fixed effects 

pFtest(et_cigs_model, cigs_model)

# p-value < 0.001 --> reject null the hypothesis that it wasn't improved 
# There is a significant additive effect of adding time fixed effects 
```




